CDSR Human Data/Modelling Paper

### 2013-09-12 scenarios/cdsr/trunk/papers/ACS2013/main.tex
JK references to situation changed to context

Section~\ref{sec:cdsrs} discusses the features of CDSRs using illustrative examples.  We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.  We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new situations, and reuse learned knowledge in multiple tasks.  To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

The long-term goal of our work is to create cognitive systems that can naturally interact with people in human environments.  While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.  First, the representational approach should be transferable to new situations.  While humans do not demonstrate mastery after a single example of a concept, they are able to begin using it immediately refining it over time.  Second, the representation should be applicable in different reasoning tasks.  The functional nature of these regions makes them important for many tasks from action planning to activity recognition.

@@

We developed the study described below in order to examine the following questions about CDSRs:
\begin{enumerate}
	\item Given a particular situation how well do people agree on the location and extent of a named CDSR?
	\item Does the orientation and location of objects (with relevant functional semantics) in a given situation effect the location and extent of a named CDSR?
\end{enumerate}

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextual dependent then one would expect a variation across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement between subjects. Research question 2 address the contextual nature of CDSRs. Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

To examine these questions, we developed two experimental tasks:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene where a particular location in the scene had been marked and an accompanying spatial description and were asked to rate how well the marked location fitted with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

\subsection{Task 1 - Results}
We use two methods to analyze the results of region drawing task.  First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous situations (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.

@@

Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e. there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel situations for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

@@

In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli. It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses. Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted situation, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded. Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject. Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$. $\kappa$ scores in this range can be interpreted as fair to good.

The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision situation $a$ had an average pairwise agreement of $0.0996$.

@@

@@

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli.

Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j). As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j). This higher rating is statiscially significant at a $95\%$ confidence level. Stimuli (k) and (l) present and analagous situation with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular situation depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted situation was consistent across similar situations, as shown by the fact that the sweetspot flipped from P1 to P3 when the situation was flipped.

@@

This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task. Applying this model to the points and regions indicated in our experiment illustrates two limitations.  Regions do not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way into the corners.  In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.  This indicates that people consider the dynamics of the situation when evaluating these terms, which is not accounted for in this model.

>>>>

Section~\ref{sec:cdsrs} discusses the features of CDSRs using illustrative examples.  We claim that (1) people can effectively communicate using these regions, and (2) the size and shape of these regions change with the context.  We argue that to account for this behavior, a cognitive systems solution must respond to changing context, apply learned regions to new contexts, and reuse learned knowledge in multiple tasks.  To explore these claims, we present the results of a human study, in which, people marked regions and rated points representing CDSRs in three different environments in 13 different contexts.  In discussing the results, we consider a number of approaches from current cognitive systems and robotics.  Each approach has a number of shortcomings that must be addressed not only to improve our understanding of human spatial reasoning, but also to enable robust intelligent assistants in human environments.

@@

The long-term goal of our work is to create cognitive systems that can naturally interact with people in human environments.  While the study presented below illustrates the importance of context, object types and configurations, on human spatial judgements, constructing a cognitive systems approach requires more than reproducing these results.  First, the representational approach should be transferable to new contexts.  While humans do not demonstrate mastery after a single example of a concept, they are able to begin using it immediately refining it over time.  Second, the representation should be applicable in different reasoning tasks.  The functional nature of these regions makes them important for many tasks from action planning to activity recognition.

@@

We developed the study described below in order to examine the following questions about CDSRs:
\begin{enumerate}
	\item Given a particular context how well do people agree on the location and extent of a named CDSR?
	\item Does the orientation and location of objects (with relevant functional semantics) in a given context effect the location and extent of a named CDSR?
\end{enumerate}

Both of these research questions address the fundamental claim that CDSRs are contextually defined. Question 1 addresses this claim by examining variation in CDSR placement within different situtations. In a context exhibiting clear contextual clues (e.g. determining a safe zone in a battle context where all information is known and battle formations are well-defined) one would expect high agreement in human identification of the CDSR. In a less well-defined context, one might expect little to no agreement between human communicators.  Research question 2 addresses the contextual nature of CDSRs in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

%Research question 1 addresses the contextual nature of CDSRs because if CDSRs are contextually dependent then one would expect a variation %across the agreement between subjects for different context: in a context that afforded the use of a particular CDSR one would expect to see a %high agreement between subject reponses, conversely in a context where a CDSR was not appropriate one might expect little or no agreement %between subjects. Research question 2 address the contextual nature of CDSRs. Research question 2 addresses the contextual nature of CDSRs %in terms of whether or not the variation in the location and extent of a CDSR systematically follow the variations  in the given context.

To examine these questions, we developed two experimental tasks:
\begin{itemize}
	\item{Task 1:} subjects were presented with a visual representation of a spatial scene plus a linguistic description containing a CDSR and were tasked with drawing a polygon on the visual stimuli such that it defined the location and extent of the CDSR in the scene (or stated there was no such region).
	\item{Task 2:} subjects were presented with a visual representation of a spatial scene with a marked location and an accompanying spatial description and were asked to rate how well the marked location fit with their understanding of the CDSR on a five point likert scale.
\end{itemize}

@@

\subsection{Task 1 - Results}
We use two methods to analyze the results of region drawing task.  First, we use interannotator agreement for two reasons: (1) if there is fair to good overall agreement, this is evidence that people change the regions they produce based systematically with context, (2) we expect less agreement in ambiguous contexts (e.g., the circle (a) and U-shaped (c) of desk arrangements as well as the turret + 1 (l) military engagement configuration).  Second, we use the extent points of the regions to assess changes between trials.  That is, as the desks are moved back from stimuli (e) to (f), we expect the maximum x-coordinate of the region to increase.

%Kate: Question here: good agreement -> evidence that people change regions with context.  I think this needs to be reworded
% but want to talk to John about it during the telecon tomorrow first

@@

Figure \ref{fig:subj-kappa} shows a histogram of the per subject average kappa scores using ten equal width intervals covering the $\kappa$ range between 0 and 1.  It is evident from this figure that most subjects had a $\kappa$ score between $0.2-0.5$ with a overall majority of subjects having a $\kappa$ score between $0.3-0.4$. A $\kappa$ score between $0.3-0.4$ would be considered as indicating poor to fair agreement, i.e. there was some agreement between subjects but less that would be expected. However, this calculation was done across all the stimuli and as we have already mentioned some of the stimuli were deliberately designed to be non-prototypical, in particular we expected the classroom stimuli where the chairs were arranged in a circle (Figure \ref{fig:classroom-a}) and in a U shaped pattern (Figure \ref{fig:classroom-c} and its horizontally flipped equivalent) to be relatively novel contexts for the subjects. Indeed, if we look at the histogram of the average pairwise $\kappa$ scores by stimuli, Figure \ref{fig:stim-kappa}, it is evident that there is a large variation, ranging from $0$ to $0.8$, in $\kappa$ scores for each stimuli.

@@

%\begin{table}
%\centerline{
%\begin{scriptsize}
%\begin{tabular}{l c c c c c c c c c c c c c}
%% (h) - 1_2; (g) - 1_1; (k) - 3_3; (f) - 2_4; (i) - 3_1; (e-flipped) - 2_3; (j) - 3_2; (f-flipped) - 2_5; (e) - 2_2;  (l) - 3_4;  (c) -  2_6;  (c-flipped) - 2_7;  (a) - 2_1
%Stim ID & (h) & (g) & (k) & (f) & (i) & (e-flipped) & (j) & (f-flipped) & (e) & (l) & (c) & (c-flipped) & (a)\\ 
%Average Pairwise $\kappa$ & 0.5034 & 0.4585 & 0.4548 & 0.3930 & 0.3906 & 0.3813 & 0.3787 & 0.3720 & 0.3629 & 0.3164 & 0.2438 & 0.2414 & 0.0996 \\ 
%No Region Responses & 0 & 1 & 0& 0& 0& 2&1&2& 1& 0&2&4& 10\\
%\hline
%\end{tabular}
%\end{scriptsize}
%}
%\label{tab:stimuli-kappa-scores}
%\caption{The average pairwise $\kappa$ scores, ordered in descending order from left to right, and the number of no region response by visual stimuli. The stimuli id are based on the labels in Figure \ref{exp-stimuli} with the horizontally flipped versions of stimuli indicated.}
%\end{table}

In more detail, Table \ref{tab:stimuli-kappa-scores} lists the average pairwise $\kappa$ score for each of the visual stimuli used in the experiment and the number of no region responses for each stimuli. It is clear from this table that the non-prototypical stimuli have the lowest $\kappa$ scores and also that there is relationship between typicality and the number of no region responses. Given this relatively large variation across stimuli, which we attribute to the prototypicality of the depicted context, we computed the average pairwise $\kappa$ scores for each subject when the non-prototypical stimuli trials were excluded. Figure \ref{fig:subj-kappa-proto} presents the histogram of the resulting average pairwise subject  $\kappa$ scores by subject. Not surpringly, with the exclusion on the non-prototypical stimuli the average pairwise $\kappa$ scores increase, with over half the subjects having an average pairwise $\kappa$ score between $0.4$ and $0.6$. $\kappa$ scores in this range can be interpreted as fair to good.

%Kate: this is a place (above) that I think we could combine/cut.  I think we can talk about the non-typical classrooms more concisely.  Let's
% say something like "some of the layouts were non-typical -> less experience with the context -> we expect lower agreement -> this is what 
% we see" then we can say that if we remove these examples, agreement on the more typical scence is good.  So suggestions are
% (1) condense discussion (2) make it clear exactly what we expect for the non typical scenes and WHY this is tied back to our claim
% about the role of context in CDSRs.  I am happy to make these changes if John agrees that I am not horribly misinterpreting his fine results

The analysis of the task 1 trials indicates that, over a range of domains and spatial regions, the responses from the subjects did agree, to a good degree. Furthermore, their results support the hypothesis that the defintion of CDSRs in a given context is effected by the location and orientation of objects in the environment. For example, focusing on the classroom scenes (stimuli, $a$, $c$, $e$, and $f$) the only difference between these trials was the layout of the chairs in the room and yet there is a substantial variation across these stimuli of the agreement between subjects on the specified spatial region, namely \textit{the front of the room}. Situation $f$ resulted in an average pairwise agreement of $0.393$ and in comparision context $a$ had an average pairwise agreement of $0.0996$.

@@

%Kate: to save some space, maybe we could talk generally about location placement and only give specifics for one set of stimuli?  Just a though

@@

As noted above, we controlled for left-right bias by flipping some of the stimuli on the horizontal axis. These flipped stimuli included the classroom stimuli (c), (e) and (f). This flipping of stimuli gave us the opportunity to compare the results of the sweetspots rankings between this pairs of stimuli trials. Given the horizontal flipping of the stimuli it was expected that the sweetspot for the stimuli should also be flipped between P1 and P3. To test whether or not the expected shift in the sweetspots occurred we ran two-tailed two sample t-tests between the sweetspots of the flipped and non-flipped stimuli pairs. In all cases the results indicated that there was no statistical difference between the rating of the sweetspot for the non-flipped stimuli and the rating of the sweetspot for the flipped stimuli using a confidence level of $95\%$. In other words, the sweetspot flipped from P1 to P3 systematically with the horizontal flipping of the stimuli. %lending support for question 2 ...

Another way of testing the relationship between the variation in context and the variation in definition of the CDSRs is to look at the differences in the ranking of the P2 point in the miliatary engagement domain between the pair of stimuli (i) and (j) and the pair of stimuli (k) and (l). In stimuli (i) the friendly tanks are positioned between the P2 point and the enemy tanks whereas in stimuli (j) the P2 point is in no-mans land. Given this change in the context one would expect a higher ranking for the P2 point in context (i) than in context (j). As the results in Table \ref{tab:point-likert-scores} show this is the case with average rating of P2  stimuli (i) being higher than that for (j). This higher rating is statiscially significant at a $95\%$ confidence level. Stimuli (k) and (l) present and analagous context with P2 being protected by an extra turret in stimuli (l) but in no-mans land in stimuli (k). Again, the average rating of P2 for stimuli (l) is higher than that of (k), although this difference is not statistically significant. Notice, that the rating of the sweetspot for the stimuli (i) and (l) remain quite high. This would indicate that the increase in the rating of the P2 points for these stimuli results from an increase in the area of the CDSR rather than a moving of the locus of the CDSR. Overall, the analysis of the ratings of the different sweetspots across the non-flipped and flipped stimuli and the variation in the ratings of the P2 points in the military domain stimuli point to the fact that both the location and the extent of a CDSR is contextually dependent.

@@

Taken together these results indicate that: (a) for all stimuli there was a location that was considered to be inside the region described by the linguistic term and that other locations were considered to be outside of this region; (b) within each domain the particular location that was considered to be the sweetspot was determined by the particular context depicted in the visual stimuli (i.e., changes in the position and orientation of objects with functional significance effected the location of the region); and (c) the impact of the changes in the depicted context was consistent across similar contexts, as shown by the fact that the sweetspot flipped from P1 to P3 when the context was flipped.

@@

This method satisfies our criteria for a cognitive systems solution for the following reasons: (1) the alignment process integrates spatial and semantic knowledge about the environment, (2) analogy enables transfer from a single example, (3) while untested, by linking conceptual knowledge to perception, regions defined by anchor points should support any symbolic reasoning task. Applying this model to the points and regions indicated in our experiment illustrates two limitations.  Regions do not always physically bounded by objects.  In the classroom examples, people frequently did not extend the front of the room all way into the corners.  In the safety example, people did not indicate that it was safe behind the lone turret in front of the others.  This indicates that people consider the dynamics of the context when evaluating these terms, which is not accounted for in this model.
